Automated Metadata Generation for Fish Specimen Image Collections

### 2021-03-26 paper/paper.tex
Probably done for the night

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Joel Pepper}
\email{jcp353@drexel.edu}
\orcid{0002-1601-8729}
\author{Jane Greenberg}
\orcid{0001-7819-5360}
\affiliation{%
  \institution{Drexel University}
  \streetaddress{3675 Market St}
  \city{Philadelphia}
  \state{Pennsylvania}
  \country{USA}
  \postcode{19104}
}
\author{Yasin Bakis}
\orcid{0001-6144-9440}
\author{Henry Bart Jr.}
\orcid{0002-5662-9444}
\affiliation{
  \institution{Tulane University}
  \city{New Orleans}
  \state{Louisiana}
  \country{USA}
}
\author{David Breen}
\orcid{0002-1376-5008}
\affiliation{%
  \institution{Drexel University}
  \streetaddress{3675 Market St}
  \city{Philadelphia}
  \state{Pennsylvania}
  \country{USA}
  \postcode{19104}
}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Pepper, Greenberg, Bakis, Bart and Breen}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
    \todo{Just throwing the Hank conference abstract in here for us to work off of and pare down} Over the last several decades advances in computing, imaging, and cyberinfrastructure have had a major impact on scientific research and discovery. One area of considerable activity is the digitization of the biological specimens that have been collected worldwide by museums and other research institutions. The scanning of these specimen collections and the placement of the resulting images into easily accessible repositories on the Internet is enabling new scientific studies based on the previously unavailable data. Unfortunately, potential scientific advances are hindered by the lack of high-quality and pertinent metadata associated with the image collections. Metadata is required to search the repositories for the imaged specimens needed for a particular study. Since the collections may each contain tens of thousands of images, producing metadata for each image via a manual process is prohibitively labor-intensive and infeasible.  Methods for automatically computing metadata from images are therefore needed to fully exploit biological image repositories for scientific discovery.

@@

Metadata is often unavailable, sparse or incorrect within specimen image repositories, but is vital for subsequent machine learning, analysis and scientific discovery. Our goal is to develop image metadata generation methods that both support the novel machine learning research underway within the BGNN project, and provide a framework for future technology developments that can be deployed by repository curators to improve and bolster the metadata they provide with their specimen images. A longer term goal is to extend the image analysis methods for computing specific quantitative features in support of specific biological investigations. For example, we are able to automatically measure the length of a fish specimen.  Associating these measurements with location and acquisition date may provide insights into the influence of habitat factors on fish development/health. Since it is prohibitively expensive for scientists to manually gather this data, we are also interested in applying our tools to images of other species stored in a variety of repositories (e.g. iDigBio). The technical challenges in achieving a broader usage of our approach mostly involve training new classifiers for different types of species, learning to segment and read annotation tags, and generalizing our classifier to find and interpret different types of rulers. This presentation will report on our current efforts to automatically generate metadata for fish specimen images and offer thoughts on how to extend these techniques for other specimen image collections.
\end{abstract}

@@

Unfortunately, potential scientific advances are hindered [by image quality] as well as [more so?] the lack of high-quality and pertinent metadata associated with the image collections. Many specimen collections include Darwin Core metadata, detailing specimen taxon, geographic location, and several other descriptive aspects. Additionally, some digitization efforts record technical metadata, detailing imaging specifications. While these types of metadata are helpful for a human examining several images at a time, they are insufficient for researchers seeking to apply computational methods to examine thousands of images to determine if, for instance, a dorsal fin of a specific fish has a different shape. Since the collections may each contain tens of thousands of images, producing metadata for each image via a manual process is prohibitively labor-intensive and infeasible. Methods for automatically computing metadata from images are therefore needed to fully exploit biological image repositories for scientific discovery.~\cite{Leipzig2021.01.28.428644}\todo{Throwing in Jeremy's paper somewhere}

@@

With an accurate mask generated, it is then necessary to check whether the bounding box needs expanded or shrunk along any edges. Expansion is done first, buy checking whether any edge intersects with any of the foreground mask pixels. If one does, it is expanded out by 1 pixel. If any edges are expanded, the whole the process, masking and expansion process is repeated to account for any changes in average intensities. Once no edges contain foreground pixels, the mask is the shrunk. This process is more efficient as all that is required is to shrink each edge by one up until the point they contain one or more foreground pixels. Once the shrinkage step is accomplished, the final mask and bounding box have been generated.
\subsection{Metadata Generation}
Including what has already been discussed, the following metadata properties are being currently being generated:

@@

For the fish \verb|length| property, it is necessary to determine the furthest points from the centroid of the fish in each direction. Since fish are normally measured in a straight line from their snout down the middle of their trunk, we project every pixel onto the major axis of the fish (as a line through its centroid) before checking the distance. This projection is done by intersecting the line through the point along the tangent of the principal axis and the aforementioned centroid--principal axis line. After checking every pixel, measuring the distance between the two furthest projected points gives the length of the fish in pixels. Multiplying this by \verb|scale| gives the fish \verb|length| in centimeters.
\subsection{Consideration}
Just gonna throw some caveats in here, might be its own section might not
\begin{itemize}
    \item Program falls back to using detectron mask and bbox for metadata generation in the event the mask generation/pixel analysis fails. Failure is defined as never finding a cohesive flood filled mask that covers more than 10\% of the detectron bbox initially. Can provide a statistic on how often this occurs.
\end{itemize}
\section{Results}
Technicians employed by our team have manually generated the 22 metadata properties deemed crucial to the overall BGNN project~\cite{Leipzig2021.01.28.428644} for a large number of INHS images. \(20,699\) total entries were created by 13 technicians that spanned \(8,398\) unique images. Of these \(8,398\) images, \(7,247\) were both not part of the \verb|detectron| training set and meet our current admissibility criteria for what \verb|detectron| and the pixel analysis process can reliably handle. We ran the metadata extraction program on these \(7,247\) images.

Our automated process currently generates 7 of the 22 core metadata properties: \verb|if_fish| (\verb|has_fish|), \verb|fish_number| (\verb|fish_count|), \verb|if_ruler| (\verb|has_ruler|), \verb|specimen_angled| (\verb|clock_value|), \verb|specimen_view| (\verb|side|), \verb|brightness| (derived from \verb|background.mean| and \verb|foreground.mean|), and \verb|if_background_uniform| (\verb|background.std|). In addition, our process also calculates bounding boxes and fish lengths in centimeters.\todo{Talk more about what we can and can't verify and how}

>>>>

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Joel Pepper}
\email{jcp353@drexel.edu}
\orcid{0002-1601-8729}
\author{Jane Greenberg}
\orcid{0001-7819-5360}
\affiliation{%
  \institution{Drexel University}
  \streetaddress{3675 Market St}
  \city{Philadelphia}
  \state{Pennsylvania}
  \country{USA}
  \postcode{19104}
}
\author{Yasin Baki\c{s}}
\orcid{0001-6144-9440}
\author{Henry Bart Jr.}
\orcid{0002-5662-9444}
\affiliation{
  \institution{Tulane University}
  \city{New Orleans}
  \state{Louisiana}
  \country{USA}
}
\author{David Breen}
\orcid{0002-1376-5008}
\affiliation{%
  \institution{Drexel University}
  \streetaddress{3675 Market St}
  \city{Philadelphia}
  \state{Pennsylvania}
  \country{USA}
  \postcode{19104}
}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Pepper, Greenberg, Baki\c{s}, Bart and Breen}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
    \todo{Just throwing the Hank conference abstract in here for us to work off of and pare down. Also, we need to decide what to do for the ``teaser image''.} Over the last several decades advances in computing, imaging, and cyberinfrastructure have had a major impact on scientific research and discovery. One area of considerable activity is the digitization of the biological specimens that have been collected worldwide by museums and other research institutions. The scanning of these specimen collections and the placement of the resulting images into easily accessible repositories on the Internet is enabling new scientific studies based on the previously unavailable data. Unfortunately, potential scientific advances are hindered by the lack of high-quality and pertinent metadata associated with the image collections. Metadata is required to search the repositories for the imaged specimens needed for a particular study. Since the collections may each contain tens of thousands of images, producing metadata for each image via a manual process is prohibitively labor-intensive and infeasible.  Methods for automatically computing metadata from images are therefore needed to fully exploit biological image repositories for scientific discovery.

@@

Metadata is often unavailable, sparse or incorrect within specimen image repositories, but is vital for subsequent machine learning, analysis and scientific discovery. Our goal is to develop image metadata generation methods that both support the novel machine learning research underway within the BGNN project, and provide a framework for future technology developments that can be deployed by repository curators to improve and bolster the metadata they provide with their specimen images. A longer term goal is to extend the image analysis methods for computing specific quantitative features in support of specific biological investigations. For example, we are able to automatically measure the length of a fish specimen.  Associating these measurements with location and acquisition date may provide insights into the influence of habitat factors on fish development/health. Since it is prohibitively expensive for scientists to manually gather this data, we are also interested in applying our tools to images of other species stored in a variety of repositories (e.g. iDigBio). The technical challenges in achieving a broader usage of our approach mostly involve training new classifiers for different types of species, learning to segment and read annotation tags, and generalizing our classifier to find and interpret different types of rulers. This presentation will report on our current efforts to automatically generate metadata for fish specimen images and offer thoughts on how to extend these techniques for other specimen image collections.\todo{Need to figure out what to actually do for this CCS stuff (\url{http://dl.acm.org/ccs.cfm}) and also the key words}
\end{abstract}

@@

Unfortunately, potential scientific advances are hindered [by image quality] as well as [more so?] the lack of high-quality and pertinent metadata associated with the image collections. Many specimen collections include Darwin Core metadata, detailing specimen taxon, geographic location, and several other descriptive aspects. Additionally, some digitization efforts record technical metadata, detailing imaging specifications. While these types of metadata are helpful for a human examining several images at a time, they are insufficient for researchers seeking to apply computational methods to examine thousands of images to determine if, for instance, a dorsal fin of a specific fish has a different shape. Since the collections may each contain tens of thousands of images, producing metadata for each image via a manual process is prohibitively labor-intensive and infeasible. Methods for automatically computing metadata from images are therefore needed to fully exploit biological image repositories for scientific discovery.

@@

With an accurate mask generated, it is then necessary to check whether the bounding box needs expanded or shrunk along any edges. Expansion is done first, buy checking whether any edge intersects with any of the foreground mask pixels. If one does, it is expanded out by 1 pixel. If any edges are expanded, the whole the process, masking and expansion process is repeated to account for any changes in average intensities. Once no edges contain foreground pixels, the mask is the shrunk. This process is more efficient as all that is required is to shrink each edge by one up until the point they contain one or more foreground pixels. Once the shrinkage step is accomplished, the final mask and bounding box have been generated.

\subsubsection{Fallback} The pixel analysis process occasionally fails. This can occur if certain flood fill operations behave unexpectedly, or if the image is too washed out or otherwise atypical for the thresholding process to work correctly. In the event this happens, the original mask and bounding box generated by \verb|detectron| is used for metadata generation.
\subsection{Metadata Generation}
Including what has already been discussed, the following metadata properties are being currently being generated:

@@

For the fish \verb|length| property, it is necessary to determine the furthest points from the centroid of the fish in each direction. Since fish are normally measured in a straight line from their snout down the middle of their trunk, we project every pixel onto the major axis of the fish (as a line through its centroid) before checking the distance. This projection is done by intersecting the line through the point along the tangent of the principal axis and the aforementioned centroid--principal axis line. After checking every pixel, measuring the distance between the two furthest projected points gives the length of the fish in pixels. Multiplying this by \verb|scale| gives the fish \verb|length| in centimeters.
\begin{comment}
\subsection{Consideration}
Just gonna throw some caveats in here, might be its own section might not
\begin{itemize}
    \item Program falls back to using detectron mask and bbox for metadata generation in the event the mask generation/pixel analysis fails. Failure is defined as never finding a cohesive flood filled mask that covers more than 10\% of the detectron bbox initially. Can provide a statistic on how often this occurs.
\end{itemize}
\end{comment}
\section{Results}
Technicians employed by our team have manually generated the 22 metadata properties deemed crucial to the overall BGNN project~\cite{Leipzig2021.01.28.428644} for a large number of INHS images. \(20,699\) total entries were created by 13 technicians that spanned \(8,398\) unique images. Of these \(8,398\) images, \(7,247\) were both not part of the \verb|detectron| training set and meet our current admissibility criteria for what \verb|detectron| and the pixel analysis process can reliably handle. We ran the metadata extraction program on these \(7,247\) images.

Our automated process currently generates 7 of the 22 core metadata properties: \verb|if_fish| (\verb|has_fish|), \verb|fish_number| (\verb|fish_count|), \verb|if_ruler| (\verb|has_ruler|), \verb|specimen_angled| (\verb|clock_value|), \verb|specimen_view| (\verb|side|), \verb|brightness| (derived from \verb|background.mean| and \verb|foreground.mean|), and \verb|if_background_uniform| (\verb|background.std|). In addition, our process also calculates bounding boxes and fish lengths in centimeters.\todo{Talk more about what we can and can't verify and how. Also, need to fit the percent of the time pixel analysis failed in here somewhere.}
